---
title: "STA 601/360 Homework 10b"
author: "Yifei Wang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document:
    highlight: tango
    theme: cerulean
---

```{r setup, message=F, warning=F, echo=F}
library(tidyverse)
require(magrittr)
require(plyr)
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = 'center')
set.seed(1)
```


## 1. O-ring failure led to the Challenger disaster. We study this phenomena by considering the temperatures at which the O-ring fails. 

```{r}
# y=1 indicates faliure
Y <- c(1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0)
X <- c(53, 57, 58, 63, 66, 67, 67, 67, 68, 69, 70, 70, 70, 70, 72, 73, 75, 75, 76, 76, 78, 79, 81)
```

## 2. What is the joint sampling distribution?


## 3. Fit a generalized linear model to the data in R. What are the odds of O-ring failure for the shuttle at 72 degrees (report to 3 decimal places)?

```{r}
logit_lm = glm(Y ~ X + 1, family = "binomial")
predict.glm(logit_lm, data.frame(X = 72), type = "link")
```


## 4. What is the acceptance probability?

```{r}
a_hat = logit_lm$coefficients[1]
b_hat = logit_lm$coefficients[2]
b0 = exp(a_hat + 0.577)
```


## 5. Implement the above Metropolis-Hastings algorithm. Start it at the MLE and run it for 10000 iterations.




```{r}
# target density function
f = function(x) {
  a = x[1]
  b = x[2]
  likelihood = exp(sum( Y*(a+b*X) - log(1+exp(a+b*X)) ))
  dab = exp(a) * exp(-exp(a)/b0) / b0
  return(likelihood * dab)
}

# density function for the proposal distribution
dprop = function(conditioned_on, x) {
  a = x[1]
  b = x[2]
  da = exp(a) * exp(-exp(a)/b0)
  db = dnorm(b, mean = b_hat, sd = sqrt(0.01))
  return(da * db)
}

# sampling from the proposal distribution
rprop <- function(x) { 
    a <- log(rexp(1, 1 / exp(a_hat + 0.577))); 
    b <- rnorm(1, b_hat, sqrt(0.1^2)); 
    return(c(a, b))
}

```

```{r}
## General code for the Metropolis-Hastings algorithm

# INPUT:
#    x0 = starting point of the chain (possibly a vector)
#     f = target density function
# dprop = density function for the proposal distribution
# rprop = sampling from the proposal distribution
#     N = Monte Carlo sample size
#     B = burn-in length

# OUTPUT:
#    x = matrix of MH samples (one in each row)
#   fx = values of the posterior density at the sampled points
# rate = proportion of times the proposal draw was accepted

mh <- function(x0, f, dprop, rprop, N, B) {

  x <- matrix(NA, N + B, length(x0))
  fx <- rep(NA, N + B)
  x[1,] <- x0
  fx[1] <- f(x0)
  ct <- 0
  for(i in 2:(N + B)) {

    u <- rprop(x[i-1,])
    fu <- f(u)
    r <- log(fu) + log(dprop(x[i-1,], u)) - log(fx[i-1]) - log(dprop(u, x[i-1,]))
    R <- min(exp(r), 1)

    if(runif(1) <= R) {

      ct <- ct + 1
      x[i,] <- u
      fx[i] <- fu

    } else {

      x[i,] <- x[i-1,]
      fx[i] <- fx[i-1]

    }

  }
  out <- list(x=x[-(1:B),], fx=fx[-(1:B)], rate=ct / (N + B))
  return(out)

}
```

```{r}
results = mh(x0 = c(a_hat, b_hat), f = f, dprop = dprop, rprop = rprop, N = 10000, B = 0)
results[3]
```


